# Data sources

```{r}
library(here)
```

This project uses three major datasets. The primary data sources are [*Federal Election Commission*](http://fec.gov/data/) (FEC) and [*The Cook Political Report*](https://cookpolitical.com/2020-national-popular-vote-tracker) (Cook Political).   

FEC gathers and processes campaign finance data and discloses information about the public funding of US presidential elections. We downloaded the Contribution Receipts data and the Campaign Disbursement data from the FEC website. Cook Political collects and verifies US election returns across the country and is the authoritative source for 2020 Presidential Election voting results. 

## U.S. County Level 2020 Presidential Election Voting Results 

```{r,include=FALSE}
knitr::opts_chunk$set(opts.label="kill_prefix")
DataDescription<- readr::read_csv(here("./data/DatasetDescription.csv"))
DataDescription_add<- readr::read_csv(here("./data/DatasetDescription_add.csv"))
```

**Dataset: 2020_US_State_Level_Presidential_Results.csv**
The dataset contains 21 columns and 61 records about the voting results of 2020 presidential election by US state. This dataset was from FEC. Here we only used 6 columns.

```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription_add, caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Dataset: 2020_US_County_Level_Presidential_Results.csv**

The dataset contains 10 columns and 3159 records about the voting results of 2020 presidential election by US county. This dataset was exported from [*Tony McGovern, 2020_US_County_Level_Presidential_Results*](https://github.com/tonmcg/US_County_Level_Election_Results_08-20/blob/master/2020_US_County_Level_Presidential_Results.csv). He scraped it from results published by from Fox News, Politico, and the New York Times.



```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[1:10, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issue with this Dataset**

The election results data was collected on Nov.22, 2020, which the results do not reflect on the latest results. The election votes counting is continuing, so we don't get the final results yet. The latest data is not yet open to public.  

## Contribution Receipts

**Dataset: contribution_bind.csv**

The contribution receipts data on [fec.gov](fec.gov) includes all receipts filed by election candidates and committees with records of contributions made by individuals, organizations, and committees. 
 
Due to a huge data volume, we limited the records exported to only those contributions with an amount over or equal to 200 dollars, received within this year, and raised by the committees of two final presidential candidates of 2020 Election. The dataset with filters applied was exported from [here](https://www.fec.gov/data/receipts/?recipient_committee_type=P&data_type=processed&committee_id=C00580100&committee_id=C00703975&two_year_transaction_period=2020&min_date=01%2F01%2F2020&max_date=11%2F03%2F2020&is_individual=true&min_amount=200).

The retrieved dataset has 79 columns and 984,824 records in total. There are 302,731 entries for Trump and 682,093 entries for Biden. 

```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[13:27, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issue with this Dataset**

1. The original dataset is extremely big with over 200 million records, but the majority of them have very small amounts on the receipts, therefore we filtered out records with amounts less than 200 dollars. As a result, the retrieved dataset might not capture information for all contributors, but at the same time it largely eliminates noise that we can focus on contributors who really made significant contributions to support their favorite candidate.  

2. The dataset contains contribution amounts of negatives and zeros, which are probably relate to unsuccessful transfers.

3. Many contributors didn't provide all the information by some reason, so there are missing values and meaningless records in some columns like name, occupation, employer, and even city and states. 

4. Since the information was gathered based on filings of committees and campaigns, the records do not comply with a unified format. For example, the contributor suffix column has records for "JR", "JR.", and "JR.," which are actually the same thing. There are also errors in the spellings of states and cities.

## The Disbursements data

**Dataset: schedule_b-2020-11-08T16_10_30.csv**

The Disbursements data on [fec.gov](fec.gov) includes the operating expenditure records of House and Senate committees, Presidential committees, and PAC and Party committees since 2003. 

For this analysis, we filtered the data to focus on the records from the presidential committees of 2020 Election, which the disbursement dates are from 01/01/2020 to 11/03/2020, to be consistent with the scope of the contribution receipt dataset. The dataset with filters applied was exported from [FEC Official Database](https://www.fec.gov/data/disbursements/?spender_committee_type=P&data_type=processed&min_date=01%2F01%2F2020&max_date=11%2F03%2F2020). The retrieved dataset has 260,587 records and 78 columns.

```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[30:39, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issue with this Dataset**

1. The raw dataset we retrieved contains 260,587 entries as of Nov 18th, 2020, but the data is still getting updated with additive filings of previous disbursements being archived. The total number of records is subject to change, but the change will be trivial compared to the size of the entire dataset. Although this will not affect our analysis or overall conclusion, it would be worth paying attention to the data incompleteness here.
