# Data sources

```{r}
library(here)
```

The primary data sources of this project are [*Federal Election Commission*](http://fec.gov/data/) (FEC) and [*Fox News*](https://cookpolitical.com/). FEC gathers and processes campaign finance data and discloses information about the public funding of US presidential elections. Fox News collects and verifies the election returns across the country and is an authoritative source for 2020 Presidential Election votes results.

We used three major datasets: the [2020 Presidential Election Results dataset](https://www.foxnews.com/elections/2020/general-results) from Fox News, as well as the Contribution Receipts dataset and the Campaign Disbursements dataset from FEC. We each was responsible for downloading one dataset from its data source. More details of the datasets are as following:


## U.S. County Level Voting Results of 2020 Presidential Election

**Dataset: 2020_US_County_Level_Presidential_Results.csv**

The dataset contains 10 columns and 3159 records covering the voting results of 2020 presidential election by US county. Since the county level data was not downloadable from the website, we exported this dataset from the github repo [* 2020_US_County_Level_Presidential_Results*](https://github.com/tonmcg/US_County_Level_Election_Results_08-20/blob/master/2020_US_County_Level_Presidential_Results.csv), which contains the latest voting data scraped by Tony McGovern from Fox News website.


```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[1:10, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issues with this dataset: **

The election results data was retrieved on Dec. 5, 2020, which the results do not reflect on the latest results. The election votes counting and post-election processing are ongoing in several states, so individual statistics we used here might change.   


## Contribution Receipts

**Dataset: contribution_bind.csv**

The contribution receipts data on [fec.gov](fec.gov) includes all receipts filed by election candidates and committees with records of contributions made by individuals, organizations, and committees. 
 
Since contributions made in the election year seem to better reflect people's support on their favorite candidate, we applied a date filter to select records with the receipt date falling between 01/01/2020 and 11/03/2020. Due to a huge data volume, we limited data exported to only the contributions with an amount over or equal to 200 dollars and raised by the committees of two final presidential candidates of 2020 Election. The filtered dataset was retrieved from [here](https://www.fec.gov/data/receipts/?recipient_committee_type=P&data_type=processed&committee_id=C00580100&committee_id=C00703975&two_year_transaction_period=2020&min_date=01%2F01%2F2020&max_date=11%2F03%2F2020&is_individual=true&min_amount=200).

The exported dataset has 79 columns and 984,824 records in total. There are 302,731 entries for **Donald J. Trump for President committee** and 682,093 entries for **Biden for President committee**. 

```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[13:27, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issues with this dataset: **

1. The original dataset is extremely big with over 200 million records, but many of them were very small contribubutions, therefore we filtered out records with amounts less than 200 dollars. As a result, the retrieved dataset might not capture information of all contributors, but at the same time, it largely eliminateds noise that we can focus on contributors who made significant contributions to support their favorite candidate.  

2. The dataset contains contribution amounts of negatives and zeros, which are probably relate to unsuccessful transfers.

3. Many contributors didn't provide all the information by some reason, so there are missing values and meaningless records in some columns like name, occupation, employer, and even city and states. 

4. Since the information was gathered based on filings of committees and campaigns, the records do not comply with a unified format. For example, the contributor suffix column has records for "JR", "JR.", and "JR.," which are actually the same thing. There are also errors in the spellings of states and cities.

## The Disbursements data

**Dataset: schedule_b-2020-11-08T16_10_30.csv**

The Disbursements data from [fec.gov](fec.gov) includes the operating expenditure records of House and Senate election committees, Presidential election committees, and PAC and Party committees since 2003. 

For this project, we filtered the data to focus on the spending of presidential election committees of 2020 US Election, which the disbursement dates are from 01/01/2020 to 11/03/2020, consistent with the scope of the contribution receipt dataset. The dataset with filters applied was exported from [FEC Official Database](https://www.fec.gov/data/disbursements/?spender_committee_type=P&data_type=processed&min_date=01%2F01%2F2020&max_date=11%2F03%2F2020). The retrieved dataset has 260,587 records and 78 columns.

```{r, echo=FALSE, results='asis'}
knitr::kable(DataDescription[30:39, ], caption = "Major columns used",
             row.names = F,font_size = 10)
```

**Issues with this dataset: **

The disbursement data is still getting updated with additive filings of previous disbursements being archived. The total number of records is subject to change, but the change will be trivial compared to the size of the entire dataset. Although this will not affect our analysis or overall conclusion, it is worth attention to the data incompleteness here.



